---
title: "Assignment 2"
author: "Yeung Sui Ping 3035871587"
date: "1/20/2022"
output: html_document
---

# Network Dataset

The dataset name is Strike, and includes the following:

IStrike.net: 24 vertices (employees), 38 edges (discussed the strike in some minimum frequency), no arcs, no loops, no line values.

Strike_groups.clu: classification of the (24) employees according to age and ethnic group: 1 - Spanish-speaking employees, 2 - young (30 or younger) English-speaking employees, 3 - old (over 30) English- speaking employees.

Strike.paj: Pajek project file containing both files.

Source: <https://sites.google.com/site/ucinetsoftware/datasets/informal-communication-within-a-sawmill-on-strike?authuser=0>

# Background

In a wood-processing facility, a new management team proposed changes to the workers' compensation package, which the workers did not accept. They started a strike, which led to a negotiation stalemate. Then, management asked an outsider to analyze the communication structure among the employees because it felt that information about the proposed changes was not effectively communicated to all employees by the union negotiators.

The outside consultant asked all employees to indicate the frequency in which they discussed the strike with each of their colleagues on a 5-point scale, ranging from 'almost never' (less then once per week) to 'very often' (several times per day). The consultant used three as a cut-off value. If at least one of two persons indicated that they discussed work with a frequency of three or more, a line between them was added to the informal communication network.

The network displays fairly stringent demarcations between groups defined on age and language. The Spanish-speaking young employees, who are of age 30 or younger, are almost disconnected from the English- speaking young employees, who communicate with no more than two of the older English-speaking employees.

All ties between groups have special backgrounds. Among the Hispanics, Alejandro is most proficient in English and Bob speaks some Spanish, which explains their tie. Bob owes Norm for getting his job and probably because of this, they developed a friendship tie. Finally, Ozzie is the father of Karl.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 
# clear the working environment
rm(list=ls())

# set up working directory
setwd("/Users/visitor/Desktop/Assignment_2_YeungSuiPing")
getwd()
```

```{r}
# load the packages
library(pacman)
p_load("igraph", "ggraph", "ggplot2", "tidyverse","RColorBrewer","data.table","network","netrankr","magrittr","stringr","visNetwork")
```

```{r}
#Reading network data from files
strike <- read.paj("strike.paj")
```

```{r}
# read paj into two data frame, links and nodes.

links <- data.frame()
for (i in 1:38){
    links <- rbind(links,c(
    strike[["networks"]][["Strike.net"]][["mel"]][[i]][["outl"]],
    strike[["networks"]][["Strike.net"]][["mel"]][[i]][["inl"]],
    strike[["networks"]][["Strike.net"]][["mel"]][[i]][["atl"]][["Strike.net"]]))
    }
 names(links) <- c("from","to","weights")
 
nodes <- data.frame()
for (i in 1:24){
  nodes <- rbind(nodes,c(i,
       strike[["networks"]][["Strike.net"]][["val"]][[i]][["vertex.names"]],
       strike[["partitions"]][["Strike_groups.clu"]][i]))
}
     
names(nodes) <- c("id","name","Strike_groups")

net <- graph_from_data_frame(d=links, vertices=nodes, directed=T) 

```

```{r}
# We'll start by adding new node and edge attributes to our dataframes. 
vis.nodes <- nodes
vis.links <- links

vis.nodes$shape  <- "dot"  
vis.nodes$shadow <- TRUE # Nodes will drop shadow
vis.nodes$group_label <- ifelse(nodes$Strike_groups==1,"Spanish-speaking employees",ifelse(nodes$Strike_groups==2,"young (30 or younger) English-speaking employees","old (over 30) English- speaking employees"))
vis.nodes$title  <-# Text on click
  paste0(vis.nodes$name,", ",vis.nodes$Strike_groups,", ",vis.nodes$group_label)
vis.nodes$label  <- vis.nodes$id # Node label
vis.nodes$size   <- 30 # Node size
vis.nodes$borderWidth <- 1 # Node border width

vis.nodes$color.background <- ifelse(nodes$Strike_groups==1,"skyblue",ifelse(nodes$Strike_groups==2,"yellow","pink"))
vis.nodes$color.border <- "black"
vis.nodes$color.highlight.background <- "orange"
vis.nodes$color.highlight.border <- "grey"

vis.links$width <- 1 # line width
vis.links$color <- "gray"    # line color  
vis.links$arrows <- "to" # arrows: 'from', 'to', or 'middle'
vis.links$smooth <- FALSE    # should the edges be curved?
vis.links$shadow <- FALSE    # edge shadow

visnet <- visNetwork(vis.nodes, vis.links,main="Informal Communication within a Sawmill on Strike Network", footer= "Assignment2_by_YeungSuiPing,Raissa") 
visnet
```
